---
title: "Homework no. 2"
author: "Dawid Pludowski"
date: "April 10, 2022"
output:
  html_document:
    df_print: paged
---

```{r message=FALSE, warning=FALSE}
library(ranger)
library(DALEX)
library(DALEXtra)
library(lime)

set.seed(123)

df <- read.csv2('./../data.csv', sep=',')
df['median_house_value'] <- lapply(df['median_house_value'], FUN = as.integer)

ranger_model <- ranger(median_house_value ~., data = df)
```

## 1. Calculating model prediction

```{r}
res <- predict(ranger_model, df[2137,])$predictions
cat(res)
```

## 2. Calculating LIME decomposition 

```{r message=FALSE}
explainer_rf <- DALEX::explain(ranger_model, 
                               data = df,  
                               y = df$median_house_value,
                               label = "random forest")

model_type.dalex_explainer <- DALEXtra::model_type.dalex_explainer
predict_model.dalex_explainer <- DALEXtra::predict_model.dalex_explainer

lime_pr <- predict_surrogate(explainer = explainer_rf, 
                             new_observation = as.data.frame(df[2137,]), 
                             n_features = 6, 
                             n_permutations = 1000,
                             type = "lime")

lime_pr
plot(lime_pr)
```

`LIME` decomposition shows that `ocean_proximity` and `total_rooms` have the greatest impact on final prediction. Explanation fit is significantly low, though.

## 3. Calculating LIME decomposition for different observation

```{r}
lime_pr <- predict_surrogate(explainer = explainer_rf, 
                             new_observation = as.data.frame(df[420,]), 
                             n_features = 6, 
                             n_permutations = 1000,
                             type = "lime")

lime_pr
plot(lime_pr)
```
As shown in previous homework, `NEAR BAY` value is supposed to have positive impact on model prediction; however, here we obtained negative impact of that value. It may be caused by the fact that in terms of `longitude` and `latitude`, houses near bay has neighbor observations only in one direction. Moreover, explanation fit is really low, which may lead to unstable explanation with that method.

In both LIME decomposition number of total rooms has similar negative impact of model prediction. There is noticeable difference between impact of `longitude` in each observation, which could be explain be the fact, that little change in distance can change `NEAR BAY` to `<1H OCEAN`, while even great change of that value cannot change `INLAND` into other value. `total_rooms` and `total_bedrooms` seem to have stable impact in neighbors of both observation, maybe because that such a values are equally important independently from other attributes of house.

In summary, we may expect that some attributes, such as `longitude` or `latitude` are unstable somewhat and other, like `total_rooms` might be much more stable.