---
title: "Homework1-XAI2022-Pawel-Fijalkowski"
author: "Paweł Fijałkowski"
date: '2022-03-30'
output:
  html_document:
    df_print: paged
---


## Initial:


As the package was removed from CRAN, we needed to download it directly from GitHub (using Rstudio Cloud, as it only worked *somehow*).

```{r, load_data, message=FALSE}
EPL_2021 <- read.csv("EPL_2021.csv")
head(EPL_2021)
```

Each row in the data set is a shot attempt. Provided features ("league", "shotType", "home_team", "away_team", ...)
are supposed to help us in predicting "xG" value which essentially is a probability that this attempt will end up with a 
positive result (a goal).

Here we select features that are important in the context of our problem and do proper transformations ((X,Y) -> (distance, angle))
```{r, message=FALSE}
library(dplyr)
data <- EPL_2021 %>% select(result, X, Y, xG, h_a, situation, shotType, home_goals, away_goals, lastAction) %>%
                 mutate(status = ifelse(result == "Goal", "1", "0")) %>%
                 mutate(distanceToGoal = sqrt((105 - (X * 105)) ^ 2 + (32.5 - (Y * 68)) ^ 2)) %>%
                 mutate(angleToGoal = abs(atan((7.32 * (105 - (X * 105))) / ((105 - (X * 105))^2 + 
                (32.5 - (Y * 68)) ^ 2 - (7.32 / 2) ^ 2)) * 180 / pi)) %>%
                 select(-X, -Y, -result)
```


## Models:

We will be predicting probability of scoring a goal (status) for each shot attempt.

```{r, message=FALSE}
library(DALEX)
library(DALEXtra)
library(ranger)
```


```{r, train_model, message = FALSE}
data_processed <- data  %>% mutate(h_a = as.factor(h_a),
                                        situation = as.factor(situation),
                                        shotType = as.factor(shotType),
                                        lastAction = as.factor(lastAction),
                                        status = as.numeric(status))

model <- ranger::ranger(status ~., data = data_processed, classification = TRUE, probability = TRUE)
```
## Explanations

#### Making prediction:

```{r obs40}
obs40 <- data_processed[40, !(colnames(data_processed) %in% c("status"))]
predict_40 <- predict(model, obs40)
ifelse(predict_40$predictions[2] <= 0.5, 0, 1)
data_processed[12, "status"]
```

Model correctly predicted that attempt #40 will not end up with the goal. 

Let's create explainer!
```{r, explainer}
explainer <- explain(model, data = data_processed, y = as.numeric(data_processed$status))
```

Break Down for observation #40

```{r predict_parts40}
plot(predict_parts(explainer = explainer, new_observation = obs40))
```

The angle to the goal contributes heavily negatively to the result (scoring goal). 

Shap on observation 40

```{r shap40}
plot(predict_parts(explainer = explainer, new_observation = obs40, type = 'shap', B = 4))
```

We clearly see that low quality of the attempt (xG) and angleToGoal contributed negatively.

We will find 2 observations in which the same feature have different effect.

```{r 37}
obs37 <- data_processed[37, !(colnames(data_processed) %in% c("status"))]
plot(predict_parts(explainer = explainer, new_observation = obs37))
```

```{r 120}
obs120 <- data_processed[120, !(colnames(data_processed) %in% c("status"))]
plot(predict_parts(explainer = explainer, new_observation = obs120))
```

As we can see "xG" value can contribute negatively(120) or positively(37) to the status variable.



