---
title: "Case Studies 2022L Lab 2"
author: "Mustafa Cavus"
date: "Mar 10, 2022"
output: html_document
---

## Installing the necessary packages

* {DALEX} for creating explainer and calling the dataset
* {tidyverse} for using glimpse()
* {caret} for splitting the dataset to train and test set
* {ranger} for traning random forest model

```{r message=FALSE, warning=FALSE, include=FALSE}
library(DALEX)
library(tidyverse)
library(caret)
library(ranger)
```

## Training the model

We will use the apartments dataset from {DALEX}

```{r}
data("apartments")
head(apartments)
```

Splitting the dataset

We can use createDataPartition() from {caret} package

```{r}
set.seed(123)
# create the sample row index for train and test set
index <- createDataPartition(apartments$m2.price, p = 0.8, list = FALSE)

# using index split to data to train and test set
train <- apartments[index,]
test  <- apartments[-index,]
```


Training a random forest model

ranger() is the fast implementation of the random forest model based on C++. By using it, you can train your model quicker than the usual one.

You need to put target and train set as arguments. Then calculate the predicted values of the model on test set and compare the values of evaluation metrics.

```{r}
set.seed(123)
ranger_model <- ranger(m2.price ~., data = train)
```


# Create an explainer

In DALEX, we need to create an explainer of the model first. 

```{r}
explainer_rf <- DALEX::explain(ranger_model, 
                               data = test[,-1],  
                               y = test$m2.price)
```

# Break-Down plot

```{r}
bd_pr <- predict_parts(explainer = explainer_rf,
                       new_observation = apartments[53,],
                       type = "break_down")

bd_pr
```

Let's plot it!

```{r}
plot(bd_pr)
```

# Shapley values

```{r}
shap_pr <- predict_parts(explainer = explainer_rf,
                         new_observation = apartments[53,],
                         type = "shap")

shap_pr
```

Let's plot it!

```{r}
plot(shap_pr)
```

The boxplots can be switch-off by using the 'show_boxplots' argument as 'FALSE'.

```{r}
plot(shap_pr, show_boxplots = FALSE) 
```

For more details see the Chapter 6 and 8 from https://ema.drwhy.ai/.

# Application

Let's try explain a specific apartment on your own with BD and Shapley values.


